WT
----

Marginal Blastozone B (leaflet, mid) vs Marginal Blastzone in C (base) region 
====================================================

```{r}
library(edgeR)
```

###Read in Data
Read in raw count data per gene.

```{r}
counts <- read.delim("sam2countsResults.tsv",row.names=1)

#check the file
head(counts)
summary(counts)
colnames(counts)
#need to convert NA to 0 counts
counts[is.na(counts)] <- 0
```

###Subset per DE expirement

I am going to start by subsetting the particular treatments I am looking at. 

```{r}
colnames(counts)
WTbmbrVScmbr <- counts[,c(32:34,42:45)]
head(WTbmbrVScmbr)
colnames(WTbmbrVScmbr)
```

```{r}
group <- c(rep("wtbmbr", 3), rep("wtcmbr", 4))
d <- DGEList(counts=WTbmbrVScmbr,group=group)
```
There are two libraries that are super small, what is going on here? Should I drop them from analaysis?
```{r}
d$samples
```

```{r}
d <- d[rowSums(cpm.d>2)>=3,] 
d <- estimateCommonDisp(d,verbose=T) 
cpm.d <- cpm(d)
d <- calcNormFactors(d)
d <- estimateCommonDisp(d)
DEtest <- exactTest(d,pair=c("wtbmbr","wtcmbr"))
head(DEtest$table)

results <- topTags(DEtest, n=Inf)
head(results)

dim(results$table)
sum(results$table$FDR<.05) # How many are DE genes?
summary(decideTestsDGE(DEtest,p.value=.05))

sig.genes <- rownames(results$table[results$table$FDR<0.05,]) 
plotSmear(d,de.tags=sig.genes)
```

Subset by all the ones with a significant score

```{r}
results.sig <- subset(tt$table, tt$table$FDR < 0.05,)
head(results.sig)
dim(results.sig)
```


What are the genes that are misexpressed?
For this we need to add some annotation

Essentially we are merging two annotations files to 1.) only sig genes 2.) all genes
```{r}

annotation1<- read.delim("../ITAG2.3_all_Arabidopsis_ITAG_annotations.tsv", header=FALSE)  #Changed to the SGN human readable annotation
colnames(annotation1) <- c("ITAG", "SGN_annotation")
annotation2<- read.delim ("../ITAG2.3_all_Arabidopsis_annotated.tsv")
annotation <- merge(annotation1,annotation2, by = "ITAG")

#Making the only significant gene table
results.sig$ITAG <- rownames(results.sig)  #change row.names to ITAG for merging
results.sig.annotated <- merge(results.sig,annotation,by = "ITAG") #This is merging to only sig genes

#Making all table

results$table$ITAG <- rownames(results$table)
results.all.annotated <- merge(results$table, annotation,by = "ITAG") #This s
```


Write table with results

```{r}
write.table(results.all.annotated,"wtbmbr_wtcmbr_DE_all.txt",sep="\t",row.names=F)
write.table(results.sig.annotated,"wtbmbr_wtcmbr_DE.txt",sep="\t",row.names=F)
```

