#Looking and visualizing individual genes: PIN1
date: June 25, 2014

The goal of this report is to look at PIN1 genes in the LCM data.

```{r, eval = false}
library(rmarkdown)
render("Solyc04g040220.Rmd", "pdf_document")
```

#Specify gene of interest
```{r}
gene1 <- "Solyc10g078370"
gene2 <- "Solyc10g080880"
gene3 <- "Solyc03g118740"
```

```{r}
library(ggplot2)
library(plyr)
library(reshape2)
library(RColorBrewer)
```

```{r}
countData <- read.csv("../data/normalized_read_count.csv")
```

Set up the dataframe for analysis and visualization
```{r}
head(countData)

#Melt Data
countData <- melt(countData)
head(countData)
colnames(countData) <- c("gene", "sample", "count")
head(countData)


#set genotype

countData$genotype <- ifelse(grepl("wt", countData$sample, ignore.case = T), "wt", 
         ifelse(grepl("tf2", countData$sample, ignore.case = T), "tf2", "unknown"))

#countData$genotype <- if(grepl("wt", countData$sample)) {
#  "wt"
#  } else {
    "tf2"
#  }

#set type

countData$tissue <- ifelse(grepl("other", countData$sample, ignore.case = T), "other", 
         ifelse(grepl("mbr", countData$sample, ignore.case = T), "mbr", "unknown"))

head(countData)

#Set Region
countData$region <- ifelse(grepl("a", countData$sample, ignore.case = T), "A", 
         ifelse(grepl("c", countData$sample, ignore.case = T), "C", "B"))
```

##Plot Data
Subsetting for genes of interest.  This is specified at the start of the script. 
```{r}
gene1Data <- subset(countData, grepl(gene1, countData$gene))
gene2Data <- subset(countData, grepl(gene2, countData$gene))
gene3Data <- subset(countData, grepl(gene3, countData$gene))

graphData <- rbind(gene1Data, gene2Data, gene3Data)
graphData$sample2  <- paste(graphData$region, graphData$tissue, sep= "")
```

##Summarize Data

```{r}
sumGraph <- ddply(graphData, c("gene", "sample2", "genotype"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N) )
```

##Plot 1 
This organizing the data to look at differences between genotype.
```{r}
ggplot(sumGraph, aes(sample2, mean, fill = genotype)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  facet_grid(.~gene) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) +
  theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, vjust=1)) 
```

##Plot 2
This organizing the data to look at differences per region. 
```{r}
ggplot(sumGraph, aes(gene, mean, fill = sample2)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  facet_grid(.~genotype) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) +
  theme(text = element_text(size=20),
        axis.text.x = element_text(angle=0, vjust=1)) 
```

#All PIN genes

```{r}
head(countData)
dim(countData)
names(countData)

countData$gene <- gsub("^(.*)[.].*", "\\1", countData$gene)
countData$gene <- gsub("^(.*)[.].*", "\\1", countData$gene)

grep("Solyc10g078370", countData$gene)

allPINs <- subset(countData, gene %in% c("Solyc10g078370", "Solyc10g080880", "Solyc03g118740", "Solyc05g008060", "Solyc07g006900", "Solyc04g007690", "Solyc06g059730", "Solyc04g056620", "Solyc02g087660", "Solyc01g068410"))

allPINs$sample2  <- paste(allPINs$region, allPINs$tissue, sep= "")
head(allPINs)
sumPINGraph <- ddply(allPINs, c("gene", "sample2", "genotype", "tissue", "region"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N) )

head(sumPINGraph)
```

```{r}

head(sumPINGraph)

ggplot(sumPINGraph, aes(gene, mean, fill = sample2)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) +
  theme_bw() +
  theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, vjust=1, face="bold", size=12),
        axis.title.y = element_text(face="bold", size=30), 
        axis.text.x  = element_text(size=16),
        axis.text.y  = element_text(size=16),
        legend.title = element_text(size=30),
        strip.text.x = element_text(size = 20)) +
  facet_grid(region~genotype)
```

WT only.
```{r}

head(sumPINGraph)

WT_heat <- subset(sumPINGraph, genotype == "wt")
WT_heat <- WT_heat[,-c(3:6,8,9)]
head(WT_heat)

colnames(WT_heat)[2] <- "region"

longWT_heat <- melt(WT_heat)
head(longWT_heat)
longWT_heat<- longWT_heat[,-3]
head(longWT_heat)

longWT_heat$gene <- factor(longWT_heat$gene, names(sort(with(longWT_heat, by(value, gene, mean)))))

#Visualization
myPalette <- colorRampPalette((brewer.pal(9, "Blues")), space="Lab")

ggplot(longWT_heat, aes(y = region, x = gene, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = myPalette(50)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_equal() +
  theme_bw() +
  theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, vjust=1, face="bold", size=12),
        axis.title.y = element_text(face="bold", size=30), 
        axis.text.x  = element_text(size=16),
        axis.text.y  = element_text(size=16),
        legend.title = element_text(size=30),
        strip.text.x = element_text(size = 20))
```


