#Looking and visualizing individual genes: PIN1
date: June 25, 2014

The goal of this report is to look at PIN1 genes in the LCM data.

```{r, eval = FALSE}
library(rmarkdown)
render("PIN1.Rmd", "pdf_document")
```

#Specify gene of interest
```{r}
gene1 <- "Solyc10g078370"
gene2 <- "Solyc10g080880"
gene3 <- "Solyc03g118740"
```

```{r}
library(reshape2)
library(ggplot2)
library(plyr)
```

```{r}
countData <- read.csv("../data/normalized_read_count.csv")
```

Set up the dataframe for analysis and visualization
```{r}
head(countData)

#Melt Data
countData <- melt(countData)
head(countData)
colnames(countData) <- c("gene", "sample", "count")
head(countData)


#set genotype

countData$genotype <- ifelse(grepl("wt", countData$sample, ignore.case = T), "wt", 
         ifelse(grepl("tf2", countData$sample, ignore.case = T), "tf2", "unknown"))

#countData$genotype <- if(grepl("wt", countData$sample)) {
#  "wt"
#  } else {
    "tf2"
#  }

#set type

countData$tissue <- ifelse(grepl("other", countData$sample, ignore.case = T), "other", 
         ifelse(grepl("mbr", countData$sample, ignore.case = T), "mbr", "unknown"))

head(countData)

#Set Region
countData$region <- ifelse(grepl("a", countData$sample, ignore.case = T), "A", 
         ifelse(grepl("c", countData$sample, ignore.case = T), "C", "B"))
```

##Plot Data
Subsetting for genes of interest.  This is specified at the start of the script. 
```{r}
gene1Data <- subset(countData, grepl(gene1, countData$gene))
gene2Data <- subset(countData, grepl(gene2, countData$gene))
gene3Data <- subset(countData, grepl(gene3, countData$gene))

graphData <- rbind(gene1Data, gene2Data, gene3Data)
graphData$sample2  <- paste(graphData$region, graphData$tissue, sep= "")
```

##Summarize Data

```{r}
sumGraph <- ddply(graphData, c("gene", "sample2", "genotype"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N) )
```

##Plot 1 
This organizing the data to look at differences between genotype.
```{r}
ggplot(sumGraph, aes(sample2, mean, fill = genotype)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  facet_grid(.~gene) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, vjust=1)) +
    scale_fill_hue(l=47)
```

##Plot 2
This organizing the data to look at differences per region. 
```{r}
ggplot(sumGraph, aes(gene, mean, fill = sample2)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  facet_grid(.~genotype) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, vjust=1)) + 
  scale_fill_hue(l=47)
```




