WT
----

Marginal Blastozone vs Other, B region (middle) attempt 1
=========================================================

```{r}
library(edgeR)
```

###Read in Data
Read in raw count data per gene.
Add checknames to FALSE because it was making the columns unique. 

```{r, results = 'hide'}
counts <- read.delim("../sam2countsResults.tsv",check.names=FALSE,row.names=1)

#check the file
head(counts)
summary(counts)
colnames(counts)
#need to convert NA to 0 counts
counts[is.na(counts)] <- 0
```

###Subset per DE expirement

I am going to start by subsetting the particular treatments I am looking at. In this case I am going to get rid of wtbmbr8 and wtbother1.4, because their count are very low and this could be the reason I am getting the errors from attempt 1.

```{r}
colnames(counts)
wtbregion <- counts[,c(32:34, 37:39)]
head(wtbregion)
colnames(wtbregion)
```

```{r}
group <- c(rep("wtbmbr", 3), rep("wtbother", 3))
d <- DGEList(counts=wtbregion,group=group)
```

```{r}
d$samples
```

```{r}
cpm.d <- cpm(d)
d <- d[rowSums(cpm.d>5)>=3,] 
d <- estimateCommonDisp(d,verbose=T) #No error this time. 

d <- calcNormFactors(d)
d <- estimateCommonDisp(d)
DEtest <- exactTest(d,pair=c("wtbmbr","wtbother"))
head(DEtest$table)

sum(DEtest$table$PValue<.05)
summary(decideTestsDGE(DEtest,p.value=.05))
sig.genes <- rownames(DEtest$table[DEtest$table$PValue<0.05,])
plotSmear(d,de.tags=sig.genes)
```
Subset by all the ones with a significant score

```{r}
results.sig <- subset(DEtest$table, DEtest$table$PValue < 0.05)
```


What are the genes that are misexpressed?
For this we need to add some annotation


```{r, results = 'hide'}
annotation1<- read.delim("../ITAG2.3_all_Arabidopsis_ITAG_annotations.tsv", header=FALSE)  #Changed to the SGN human readable annotation
colnames(annotation1)<- c("ITAG", "SGN_annotation")
annotation2<- read.delim ("../ITAG2.3_all_Arabidopsis_annotated.tsv")
annotation <- merge (annotation1,annotation2, by =1,1, all.x=TRUE)
head(annotation) 
results.annotated <- merge(results.sig,annotation,by.x="row.names",by.y="ITAG",all.x=T,sort=F)
```

Write table with results
```{r}
write.table(results.annotated,"wtbmbr_wtbother_DE1.attempt2.txt",sep="\t",row.names=F)
```

