WT
----

Marginal Blastozone A vs Other, A region (tip)
=========================================================

```{r}
library(edgeR)
```

###Read in Data
Read in raw count data per gene.
Add checknames to FALSE because it was making the columns unique. 

```{r, results='hide'}
counts <- read.delim("../sam2countsResults.tsv",check.names=FALSE,row.names=1)

#check the file
head(counts)
summary(counts)
colnames(counts)
#need to convert NA to 0 counts
counts[is.na(counts)] <- 0
```

###Subset per DE expirement

I am going to start by subsetting the particular treatments I am looking at.

```{r}
colnames(counts)
wtaregion <- counts[,c(24:26, 27:31)]
head(wtaregion)
colnames(wtaregion)
```

```{r}
group <- c(rep("wtambr", 3), rep("wtaother", 5))
d <- DGEList(counts=wtaregion,group=group)
```

Here are all the samples. Why is the lib.size NA?
```{r}
d$samples
```

```{r}
cpm.d<- cpm(d)
d <- d[rowSums(cpm.d>5)>=3,] 
d <- estimateCommonDisp(d,verbose=T) 
d <- calcNormFactors(d)
d <- estimateCommonDisp(d)
DEtest <- exactTest(d,pair=c("wtambr","wtaother"))

tt <- topTags(DEtest, n=Inf)
head(tt)
sum(DEtest$table$PValue<.05)
summary(decideTestsDGE(DEtest,p.value=.05))
sig.genes <- rownames(DEtest$table[DEtest$table$PValue<0.05,])
plotSmear(d,de.tags=sig.genes)
```
Subset by all the ones with a significant score

```{r}
results.sig <- subset(DEtest$table, DEtest$table$PValue < 0.05)
```


What are the genes that are misexpressed?
For this we need to add some annotation


```{r, results='hide'}
annotation1<- read.delim("../ITAG2.3_all_Arabidopsis_ITAG_annotations.tsv", header=FALSE)  #Changed to the SGN human readable annotation
colnames(annotation1)<- c("ITAG", "SGN_annotation")
annotation2<- read.delim ("../ITAG2.3_all_Arabidopsis_annotated.tsv")
annotation <- merge (annotation1,annotation2, by =1,1, all.x=TRUE)
head(annotation) 
results.annotated <- merge(results.sig,annotation,by.x="row.names",by.y="ITAG",all.x=T,sort=F)
```

Write table with results
```{r}
write.table(results.annotated,"wtambr_wtaother.txt",sep="\t", row.names=F)
```

