#Looking and visualizing individual genes: PIN1
date: June 25, 2014

The goal of this report is to look at PIN1 genes in the LCM data.

```{r, eval = FALSE}
library(rmarkdown)
render("curatedGenes.Visualized.Rmd", "pdf_document")
```

```{r}
library(reshape2)
library(ggplot2)
library(plyr)
```

##Read in Data
```{r}
countData <- read.csv("../../data/normalized_read_count.csv")
geneList1 <- read.csv("pnas.1402835111.sd06.csv")
geneList2 <- read.csv("pnas.1402835111.sd07.csv")
```

Set up the counts dataframe for analysis and visualization
```{r}

#Melt Data
countData <- melt(countData)
colnames(countData) <- c("gene", "sample", "count")


#set genotype

countData$genotype <- ifelse(grepl("wt", countData$sample, ignore.case = T), "wt", 
         ifelse(grepl("tf2", countData$sample, ignore.case = T), "tf2", "unknown"))

#set tissue

countData$tissue <- ifelse(grepl("other", countData$sample, ignore.case = T), "other", 
         ifelse(grepl("mbr", countData$sample, ignore.case = T), "mbr", "unknown"))

#Set Region
countData$region <- ifelse(grepl("a", countData$sample, ignore.case = T), "A", 
         ifelse(grepl("c", countData$sample, ignore.case = T), "C", "B"))

#Set type

countData$type <- paste(countData$region, countData$tissue,  sep = "")

```


##Subset for each gene

First I take the genes and make them into a characters, so that I can find them and loop through. Takes ~ 20 minutes.

```{r}
genes2 <- geneList1[,1]
genes1 <- geneList2[,1]

genesOfInterest1 <- levels(genes1)
genesOfInterest2 <- levels(genes2)

genesOfInterest <- c(genesOfInterest1, genesOfInterest2)
```

Intitialize datafram to fit in all the visualization information. 

```{r}
largeGeneList <- data.frame(t(rep(NA,7)))
colnames(largeGeneList) <- c("type", "genotype", "N", "mean", "sd", "se", "gene")
head(largeGeneList)
```
##Making the data table for Visualization 

This loop 
(Takes about 20 minutes) 

```{r, tidy = FALSE}
for(GENE in genesOfInterest) {
 
  if(length(grep(GENE, countData$gene)) < 1){ #this is just making sure that the curated
    next; 
    }
      
  geneData <- subset(countData, grepl(GENE, countData$gene))
   
  sumGraph <- ddply(geneData, c("type", "genotype"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N))
  
  sumGraph$gene <- GENE
  
largeGeneList  <- rbind(largeGeneList, sumGraph) #bind together all the new rows per loop. 
}

#make for use
finalList <- largeGeneList
finalList <- finalList[-1,] #remove the first row

#get the log2 of mean. 

finalList$log2Mean <- log2(finalList$mean) # why are there negative numbers now?
finalList <- subset(finalList, log2Mean > 0 ) #fix that problem, but must go back to figure out why

#Is order for the lines to correct properly, they must be grouped by both genotype and gene.  So I am going to combine columns gene and genotype for the proper grouping. 

finalList <- within(finalList, lineGroup <- paste(genotype, gene,sep='.'))
head(finalList)

write.table(finalList, file = "../../data/allGeneListCurated.csv", sep = ",")
```

#Visualization

Now that I have the dataset, I can begin visualization. the main ways I want to visualize are:

1. All genes, all tissue types, colored by genotype. 

```{r}

ggplot(finalList, aes(type, log2Mean, group = lineGroup, color =  genotype )) + 
  geom_line(alpha = .1, (aes(color = factor(genotype)))) + 
  geom_point(alpha = .0)
```

